---
title: "Trichodesmium_metaproteome_code"
author: "Hanna Anderson"
date: "2025-01-07"
output:
  pdf_document: default
  html_document: default
---
# Code for figures, intermediate data products, and statistical analysis for Anderson et al. (2025) 

# packages needed
```{r}
library(rlang)
library(tibble)
library(tidyr)
library(dplyr)
library(ggVennDiagram)
library(vegan)
library(ggforce)
library(pheatmap)
library(reshape2)
library(ggpubr)
library(rnaturalearth)
library(sf)
library(Hmisc)
library(corrplot)
```

# read in Table S1 as a .csv file for the metadata df
```{r}
#setwd("") # set your working directory (wd)
env_initial <- read.csv("S1.csv", na.strings = "")

env <- env_initial %>% 
  filter(row_number() <= n()-3)  %>% # remove 3 empty rows at bottom of file
  arrange(Ocean_region) # order samples alphabetically by region
```

# read in Table S3 as a .csv file for the protein relative abundance df
```{r}
#setwd("") # set your ed
tricho_initial <- read.csv("S3.csv")
```

# set up Trichodesmium protein relative abundance df
```{r}
tricho_2 <- tricho_initial %>%  
  mutate(Atlantic_N_count = rowSums(tricho_initial[, c(4:17)], na.rm=TRUE)) %>%
  mutate(Pacific_N_count = rowSums(tricho_initial[, c(18:26)], na.rm=TRUE)) %>%
  mutate(Pacific_S_count = rowSums(tricho_initial[, c(27:29)], na.rm=TRUE)) %>%
  mutate(Avg_N_Atl = (Atlantic_N_count / 14)) %>%
  mutate(Avg_N_Pac = (Pacific_N_count / 9)) %>%
  mutate(Avg_S_Pac = (Pacific_S_count / 3))

# apply cutoff to dataset (remove proteins with <1 spectra in any of the 3 regions)
tricho <- tricho_2[tricho_2$Avg_N_Atl > 1 | tricho_2$Avg_N_Pac > 1  | tricho_2$Avg_S_Pac > 1,]
```

# figure 1b venn diagram code
```{r}
# subset tricho df to tricho and epibiont proteins
t <- tricho[tricho$Who == "tricho" ,]

# create vectors of unique annotations for each basin
Atl_N_tricho <- t[!t$Atlantic_N_count == 0, "Annot"]
Pac_N_tricho <- t[!t$Pacific_N_count == 0 ,"Annot"]
Pac_S_tricho <- t[!t$Pacific_S_count == 0 ,"Annot"]

# put the vectors into a list 
input_tricho <- list(Atl_N_tricho=Atl_N_tricho, Pac_N_tricho=Pac_N_tricho, Pac_S_tricho=Pac_S_tricho)

# create venn diagram
venn <- ggVennDiagram(input_tricho, edge_size = 1, category.names = c("North \nAtlantic","North Pacific","South \nPacific"), label_alpha = 0, set_size = 7, label_size = 7, set_color = "black") +
scale_fill_gradient(low = "#e6e6e6", high = "#666666") +  
  labs(fill="Number of \nProteins") +
  scale_x_continuous(expand = expansion(mult = .2)) 
```

# figure 1a map code 
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf") # return world country polygons
class(world) # verify what you made above

# define a map projection
projcrs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# convert lon and lat from env df into an sf object
df <- st_as_sf(x = env,                         
           coords = c("Lon", "Lat"),
           crs = projcrs)

# assign Ocean_region column as.factor for control levels
df$Ocean_region <- as.factor(df$Ocean_region)
levels(df$Ocean_region) <- c("North Atlantic","North Pacific","South Pacific")

# pick colors for pretty figures (optional)
colors <- c(rep('#55C667',14), rep("#97DFFC",9), rep("#4A8FE7",3))
# fill = "darkgrey" for no color figure

map <- ggplot(data = world) +
    geom_sf(fill="gray76") +
    theme(panel.grid = element_line(linetype = "dashed", 
        size = 0.5, color ="grey76"), panel.background = element_rect(fill="grey96"),legend.text = element_text(size=20), legend.title = element_text(size=20), axis.text = element_text(size=20)) +
    geom_sf(data = df, aes(shape=Ocean_region), color="black", size=5, stroke=1.5, fill = colors) +
        scale_shape_manual(values = c(21, 22, 23)) +
    coord_sf(xlim = c(-170.15, -45.12), ylim = c(-30, 33.97), expand = TRUE) +
    labs(shape="Ocean Basin", fill = colors) +
  theme(legend.position = c(0.4, 0.25),  legend.background = element_rect(linetype = 1, linewidth= 0.5, colour = 1), legend.title.align = 0.5)
```

# figure 2a and S1 CA plot code
```{r}
# remove all epibiont proteins
ca_proteins <- tricho[!tricho$Who == "epi",]

# to generate Fig. S1, make the CA with only proteins found in all basins. add the following line of code:
#ca_proteins <- ca_proteins[!ca_proteins$Avg_N_Atl == 0 & !ca_proteins$Avg_N_Pac == 0 & !ca_proteins$Avg_S_Pac == 0,]

# prepare df for ordination analysis
ca_proteins_2 <- ca_proteins %>%
  separate(Annot, c(NA,'Uniprot.ID'), sep = '_', remove = FALSE) %>%
  remove_rownames %>% 
  column_to_rownames(var="Uniprot.ID") %>%
  dplyr::select(-c(1:3,30:56)) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame()

ca_proteins_3 <- ca_proteins_2[order(row.names(ca_proteins_2)),] # order by row names

# make sure env df is edited for vector analysis
ca_env <- env[order(env$Label_by_location),]

ca_env_2 <- ca_env %>%
  dplyr::select(-c(12:13)) %>%
  remove_rownames() %>%
  column_to_rownames('Label_by_location') %>%
  dplyr::rename(dFe = DFe_nM) %>%
  dplyr::rename(DIP = DIP_nM) %>%
  dplyr::rename(N_N = N.N_nM)

# ordination for protein relative abundance with vegan package
ord <- cca(ca_proteins_3)

# envfit on environmental variables
fit <- envfit(ord ~ dFe + DIP + Temp_C, ca_env_2, perm = 9999, na.rm = TRUE)

# putting vegan objects into ggplot2 objects
scrs <- as.data.frame(scores(ord, display="sites"))
scrs <- cbind(scrs, Basin = c(rep("North Atlantic", 14), rep("North Pacific",9), rep("South Pacific",3)))

# environmental vectors as ggplot objects
spp.scrs <- as.data.frame(scores(fit, display = "vectors"))
spp.scrs <- cbind(spp.scrs, Var = rownames(spp.scrs))

shapes <- c(rep(21, 14), rep(22,9), rep(23,3)) #assign shapes
colors_2 <- c(rep('#55C667',14), rep("#97DFFC",9), rep("#4A8FE7",3)) #assign colors, if you want

# make plot
ca <- ggplot(scrs) +
  geom_point(mapping = aes(x=CA1, y=CA2), color="black",fill="grey", stroke=1.5, size=7.5, shape=shapes) +
  coord_fixed() +
  geom_segment(data=spp.scrs, aes(x=0, xend=CA1*2, y=0, yend=CA2*2), arrow=arrow(length=unit(0.65, "cm")), color="darkgrey", linewidth=1) +
  geom_text(data=spp.scrs, aes(x= CA1*2, y = CA2*2, label=Var), size=5, position=position_dodge(width=1), vjust=-0.5) +
  geom_mark_ellipse(data=scrs, aes(x=CA1, y=CA2, color=colors_2, label=Basin), expand = unit(0.5, "mm"), size=0, fill = "grey") +
  geom_hline(yintercept=0, color="grey", linetype="dashed") +
  geom_vline(xintercept=0, color="grey", linetype="dashed") +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), axis.title = element_text(size=20), axis.text = element_text(size=18) ) +
scale_colour_manual(breaks = unique(scrs$Basin),
                    values = shapes) +
  guides(color=FALSE) +
  labs(title="CA analysis with proteins from all basins")
```

# figure 2b heatmap code
```{r}
# read in Table S7 as a .csv file for the metadata df
# Column submodule assignment has been modified according to rules in manuscript SI, original column is also in the df

#setwd("") #set your wd
heatmap <- read.csv("S7.csv", na.strings = "")

# make sure epibiont proteins are removed
heatmap_1 <- heatmap %>%
  filter(Who == "tricho")

# prepare df for analysis
heatmap_2 <- heatmap_1 %>%
  dplyr::mutate(Atlantic_N_count = rowSums(heatmap_1[, c(3:16)], na.rm=TRUE)) %>%
  dplyr::mutate(Pacific_N_count = rowSums(heatmap_1[, c(17:25)], na.rm=TRUE)) %>%
  dplyr::mutate(Pacific_S_count = rowSums(heatmap_1[, c(26:28)], na.rm=TRUE)) %>%
  dplyr::mutate(Avg_N_Atl = (Atlantic_N_count / 14)) %>%
  dplyr::mutate(Avg_N_Pac = (Pacific_N_count / 9)) %>%
  dplyr::mutate(Avg_S_Pac = (Pacific_S_count / 3)) %>%
  dplyr::select(c(35,49:51)) %>% # only keep annotations and basin avgs
  group_by(Submodule_Assignment) %>% 
  mutate_if(is.numeric, sum) %>% # sum submodule assignment relative abundances
  distinct() %>%
  ungroup() 

# order the heatmap by KEGG submodule label
heatmap_3 <- heatmap_2[order(heatmap_2$Submodule_Assignment),]

# for basins, percent of counts per the total
heat <- heatmap_3 %>% 
  dplyr::mutate(percent_N_Atl = round(Avg_N_Atl/sum(Avg_N_Atl),7)) %>%
  dplyr::mutate(percent_N_Pac = round(Avg_N_Pac/sum(Avg_N_Pac),7)) %>%
  dplyr::mutate(percent_S_Pac = round(Avg_S_Pac/sum(Avg_S_Pac),7)) %>%
  dplyr::select(-c(2:4))

# log transform df
heat[2:4] <- lapply(heat[2:4], log)

# change -Inf to NA
heat[heat == "-Inf"] <- NA 

# remove low-abundance or irrelevant submodules
heat_1 <- heat[!heat$Submodule_Assignment == "Aging" & 
                 !heat$Submodule_Assignment == "Biosynthesis of secondary metabolites" & 
                 !heat$Submodule_Assignment == "Chaperones and folding catalysts" & 
                 !heat$Submodule_Assignment == "Chromosome and associated proteins" & 
                 !heat$Submodule_Assignment == "Cytochrome P450" & 
                 !heat$Submodule_Assignment == "Glycosyltransferases" & 
                 !heat$Submodule_Assignment == "Human Diseases" & 
                 !heat$Submodule_Assignment == "Membrane transport" & 
                 !heat$Submodule_Assignment == "Mitochondrial biogenesis" & 
                 !heat$Submodule_Assignment == "Organismal systems" & 
                 !heat$Submodule_Assignment == "Organismal Systems" & 
                 !heat$Submodule_Assignment == "Signaling and cellular processes" & 
                 !heat$Submodule_Assignment == "Xenobiotics biodegradation and metabolism" &
                 !heat$Submodule_Assignment == "ABC-2 type and other transport systems" &
                 !heat$Submodule_Assignment == "Protein kinases" &
                 !heat$Submodule_Assignment == "Polyketide biosynthesis proteins" &
                 !heat$Submodule_Assignment == "Glycine, serine and threonine metabolism", ]

# make lists to rename columns
oldnames = c("percent_N_Atl","percent_N_Pac","percent_S_Pac")
newnames = c("North Atlantic", "North Pacific", "South Pacific")

# final prep before heatmap plotting
heater <- heat_1 %>% 
  rowwise() %>%
  dplyr::mutate(avg = mean(c_across(c('percent_N_Atl', 'percent_N_Pac', 'percent_S_Pac')), na.rm=TRUE)) %>% # create a column to arrange submodules by average descending percent
  dplyr::arrange(desc(avg)) %>% 
  column_to_rownames(var = 'Submodule_Assignment') %>%
  dplyr::select(-c(4)) %>% # remove avg column 
  rename_at(vars(oldnames), ~ newnames) %>% 
  as.matrix()

# submodules heatmap made with to include breaks, organized by module
my_heatmap <- pheatmap(heater, color=turbo(50), legend=T, 
                      cluster_rows = FALSE, cluster_cols=FALSE, 
                      na_col="white", fontsize=12)
```

# figure 3
```{r}
# filter for Fe and P marker proteins (Table S4) that are found in all basins
boxplot_norm <- tricho %>%
  filter(Who == "tricho")

# change 0 values to NA
boxplot_norm[boxplot_norm==0] <- NA

# calculate averages across rows for regions
boxplot_norm_1 <- boxplot_norm %>%
  mutate(Avg_N_Atl1 = rowMeans(boxplot_norm[4:17], na.rm=TRUE)) %>%
  mutate(Avg_N_Pac1 = rowMeans(boxplot_norm[18:26], na.rm=TRUE)) %>%
  mutate(Avg_S_Pac1 = rowMeans(boxplot_norm[27:29], na.rm=TRUE)) %>%
  mutate(Avg_Basins1 = rowMeans(boxplot_norm[ , c(4:29)], na.rm=TRUE))

# normalize regional averages by the average of all basins and add a category for marker proteins
boxplot_norm_2 <- boxplot_norm_1 %>% 
  mutate(boxplot_norm_1, Norm_N_Atl1 = Avg_N_Atl1/Avg_Basins1) %>%
  mutate(boxplot_norm_1, Norm_N_Pac1 = Avg_N_Pac1/Avg_Basins1) %>%
  mutate(boxplot_norm_1, Norm_S_Pac1 = Avg_S_Pac1/Avg_Basins1) %>%
  mutate(Category = if_else(Tery_ID == "Tery_1667" | Tery_ID == "Tery_3377" |
                            Tery_ID == "Tery_1666" | Tery_ID == "Tery_1687",
                            "Iron Stress",
                    if_else(Tery_ID == "Tery_5003" | Tery_ID ==
                            "Tery_4993" | Tery_ID == "Tery_3845" |
                            Tery_ID == "Tery_3534" |  Tery_ID == "Tery_0398", 
                            "Phosphorus Stress", NA))) %>%
  filter(!(is.na(Category))) %>% # keep only marker proteins
   dplyr::select(c(61:64)) # keep category and region average columns

# melt df
boxplot_melt <- boxplot_norm_2 %>%
  reshape2::melt(id.var= c(4), variable.name = 'Basin')

# assign region names to Basin column
levels(boxplot_melt$Basin) <- c("North Atlantic","North Pacific","South Pacific")

# make sure Category and Basin are characters
boxplot_melt$Category <- as.character(boxplot_melt$Category)
boxplot_melt$Basin <- as.character(boxplot_melt$Basin)

# isolate phosphorus stress markers
phos <- as.data.frame(boxplot_melt[boxplot_melt$Category == "Phosphorus Stress",])
phos$Basin <- as.factor(phos$Basin)

# isolate iron stress markers
fe <- as.data.frame(boxplot_melt[boxplot_melt$Category == "Iron Stress",])
fe$Basin <- as.factor(fe$Basin)

# make color lists
colors <- c('#4D4D4D', "#AEAEAE", "#E6E6E6")
colors_2 <- c('#55C667', "#97DFFC", "#4A8FE7") # for colorful figures, if you want

# generate the plot with code below
par(mfrow=c(1,2))

par(mar = c(7, 5, 5, 5) + 0.3)   # P stress            
boxplot(phos$value ~ phos$Basin, main="Phosphorus Stress Markers", xlab = "", ylab = "P Marker Protein Stress Signal", cex.lab=2, cex.axis=1.5, cex.main=2, cex.sub=1.6, col=colors)
abline(h=1, lty=2)

par(mar = c(7, 5, 5, 5) + 0.3)   # Fe stress           
boxplot(fe$value ~ fe$Basin, xlab = "", ylab = "Fe Marker Protein Stress Signal",  main="Iron Stress Markers", cex.lab=2, cex.axis=1.5, cex.main=2, cex.sub=1.6, yaxt="n",  col=colors)
abline(h=1, lty=2)
axis(side = 2, at = c(0.5,1,1.5,2), cex.axis=1.7)
```

# figure 4 code
```{r}
# plotting Fe and P stress signal by sample can be done by reading in Table S6
# this data can also be generated from Table S3, the df we have been using thus far (using code below)

# normalize all sample columns by average of all samples
stress_1 <- tricho %>%
  mutate(across(is.numeric, ~na_if(., 0))) %>% # turn zeros to NA
  mutate(Avg_Basins1 = rowMeans(tricho[ , c(4:29)], na.rm=TRUE)) %>% # create a column which averages all proteins across samples
  replace(is.na(.), 0) %>% # turn NAs back to zeros
  mutate(across(c(4:29), ~ . / Avg_Basins1, .names = "{.col}_norm")) # normalize proteins by dividing each column by the average column

# subset to stress markers and take averages across marker sets 
stress_by_sample <- stress_1 %>%  
  mutate(Category = if_else(Tery_ID == "Tery_1667" | Tery_ID == "Tery_3377" |
                            Tery_ID == "Tery_1666" | Tery_ID == "Tery_1687",
                            "Iron_Stress",
                    if_else(Tery_ID == "Tery_5003" | Tery_ID ==
                            "Tery_4993" | Tery_ID == "Tery_3845" | 
                            Tery_ID == "Tery_3534" |  Tery_ID == "Tery_0398"
                           , "Phosphorus_Stress", NA))) %>%
  filter(!(is.na(Category))) %>% # keep only marker proteins
  dplyr::select(c(58:84)) %>%
  mutate(across(is.numeric, ~na_if(., 0))) %>% # turn zeros to NA
  group_by(Category) %>%
  summarise(across(everything(), mean, na.rm = TRUE)) %>% # take a mean of all Fe or P marker proteins for each sample
  replace(is.na(.), 0) %>% # turn NAs back to zeros
  ungroup() %>%
  remove_rownames() %>%
  column_to_rownames("Category") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Station")

# trim characters from Station string to match other dataset for left_join
stress_by_sample$Station <- substr(stress_by_sample$Station, 1, nchar(stress_by_sample$Station)-5) 

 # select a few proteins out
select_few <- stress_1 %>%
  filter(Tery_ID == "Tery_3845" | Tery_ID == "Tery_3467" | Tery_ID == "Tery_4136") %>%
  dplyr::select(c(2,58:83)) %>%
  remove_rownames() %>%
  column_to_rownames("Tery_ID") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Station") %>%
  mutate(across(is.numeric, ~na_if(., 0))) %>% # turn zeros to NA
  mutate(across(c(2:6), ~ . / max(., na.rm = TRUE))) %>% # divide by the maximum number in each column to put all on a scale of 0-1
  replace(is.na(.), 0) # turn the NAs back to zeros

# trim characters from Station string to match other dataset for left_join
select_few$Station <- substr(select_few$Station, 1, nchar(select_few$Station)-5)

# join select_few proteins with marker protein stress metric
stressed <- stress_by_sample %>%
  left_join(select_few, by = "Station")

# add a caolumn for the Basin category
stressed <- cbind(stressed, Basin = c(rep("North Atlantic", 14), rep("North Pacific",9), rep("South Pacific",3)))

# generate plots for figs 4 and S2
phox <- ggplot(stressed, aes(x=Iron_Stress, y=Phosphorus_Stress, shape = Basin, fill = Tery_3845)) +
  geom_point(aes(size = 10, fill = Tery_3845))  +
  scale_fill_distiller(type = "seq", direction = 1, palette = "Greys")  +
  scale_shape_manual(values = c(21, 22, 23)) +
  geom_hline(yintercept = 0.8575063, linetype="dashed") +
  geom_vline(xintercept = 0.8732895, linetype="dashed") +
  labs(y="Average P stress marker signal", x="Average Fe stress marker signal", fill=NULL, title =NULL, x=NULL) +
    theme_bw() +
  theme(legend.box="horizontal", plot.title = element_text(hjust = 0.5, size=15), axis.title.x = element_text(size=15), axis.title.y = element_text(size=15), axis.text = element_text(size=12), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position = c(0.8,0.8), legend.background = element_rect(linetype='solid', color='grey'), legend.title.align = 0.5, legend.title = element_text(size=12), legend.text = element_text(size=12), axis.text.x=element_text(color="black"), axis.text.y=element_text(color="black")) +
  guides(size="none", shape = guide_legend(override.aes = list(size=5))) +
  labs(fill='PhoX')

phoa <- ggplot(stressed, aes(x=Iron_Stress, y=Phosphorus_Stress, shape = Basin, fill = Tery_3467)) +
  geom_point(aes(size = 10, fill = Tery_3467))  +
  scale_fill_distiller(type = "seq", direction = 1, palette = "Greys")  +
  scale_shape_manual(values = c(21, 22, 23)) +
  geom_hline(yintercept = 0.8575063, linetype="dashed") +
  geom_vline(xintercept = 0.8732895, linetype="dashed") +
  labs(y="Average P stress marker signal", x="Average Fe stress marker signal", fill=NULL, title =NULL, x=NULL) +
    theme_bw() +
  theme(legend.box="horizontal", plot.title = element_text(hjust = 0.5, size=15), axis.title.x = element_text(size=15), axis.title.y = element_text(size=15), axis.text = element_text(size=12), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position = c(0.8,0.8), legend.background = element_rect(linetype='solid', color='grey'), legend.title.align = 0.5, legend.title = element_text(size=12), legend.text = element_text(size=12), axis.text.x=element_text(color="black"), axis.text.y=element_text(color="black")) +
  guides(size="none", shape = "none") +
  labs(fill='PhoA')

nifh <- ggplot(stressed, aes(x=Iron_Stress, y=Phosphorus_Stress, shape = Basin, fill = Tery_4136)) +
  geom_point(aes(size = 10, fill = Tery_4136))  +
  scale_fill_distiller(type = "seq", direction = 1, palette = "Greys")  +
  scale_shape_manual(values = c(21, 22, 23)) +
  geom_hline(yintercept = 0.8575063, linetype="dashed") +
  geom_vline(xintercept = 0.8732895, linetype="dashed") +
  labs(y="Average P stress marker signal", x="Average Fe stress marker signal", fill=NULL, title =NULL, x=NULL) +
    theme_bw() +
  theme(legend.box="horizontal", plot.title = element_text(hjust = 0.5, size=15), axis.title.x = element_text(size=15), axis.title.y = element_text(size=15), axis.text = element_text(size=12), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.position = c(0.8,0.8), legend.background = element_rect(linetype='solid', color='grey'), legend.title.align = 0.5, legend.title = element_text(size=12), legend.text = element_text(size=12), axis.text.x=element_text(color="black"), axis.text.y=element_text(color="black")) +
  guides(size="none", shape = "none") +
  labs(fill='NifH')

f4 <- ggarrange(phox, phoa, nifh, nrow=3, labels = c("a","b","c"))
```

# figure S2
```{r}
# filter for putative Fe-P co-stress marker proteins (Table S5) that are found in all regions
costress <- tricho %>%
  filter(Who == "tricho")

# change 0 values to NA
costress[costress==0] <- NA

# calculate averages across rows for regions
costress_1 <- costress %>%
  mutate(Avg_N_Atl1 = rowMeans(costress[, c(4:17)], na.rm=TRUE)) %>%
  mutate(Avg_N_Pac1 = rowMeans(costress[, c(18:26)], na.rm=TRUE)) %>%
  mutate(Avg_S_Pac1 = rowMeans(costress[, c(27:29)], na.rm=TRUE)) %>%
  mutate(Avg_Basins1 = rowMeans(costress[ , c(4:29)], na.rm=TRUE))

# normalize regional averages by the average of all basins and add a category for marker proteins
costress_norm <- costress_1 %>% 
  mutate(costress_1, Norm_N_Atl1 = Avg_N_Atl1/Avg_Basins1) %>%
  mutate(costress_1, Norm_N_Pac1 = Avg_N_Pac1/Avg_Basins1) %>%
  mutate(costress_1, Norm_S_Pac1 = Avg_S_Pac1/Avg_Basins1)

# filter for 18 putative co-stress proteins (from Walworth et al., 2016; SUpplemental Data 3) that are found in all 3 regions in this dataset
costress_norm_1 <- costress_norm[grepl("Q10WA6|Q10WH0|Q10YN0|Q10ZM3|Q111K5|Q112L9|Q113J5|Q114G2|Q115F6|Q116D0|Q116X2|Q116X7|Q118F5|Q118N0|Q119V1|Q119Y0|Q11A61|Q11AA1", costress_norm$Annot, ignore.case = TRUE),]

costress_norm_2 <- costress_norm_1 %>%
   mutate(Category = "Walworth co-stress") %>%
   dplyr::select(c(61:64)) # keep category and region average columns

# melt df
costress_melt <- costress_norm_2 %>%
  reshape2::melt(id.var= c(4), variable.name = 'Basin')

# assign region names to Basin column
levels(costress_melt$Basin) <- c("North Atlantic","North Pacific","South Pacific")

# make sure Category and Basin are characters
costress_melt$Category <- as.character(costress_melt$Category)
costress_melt$Basin <- as.character(costress_melt$Basin)

# isolate phosphorus stress markers
wa <- as.data.frame(costress_melt[costress_melt$Category == "Walworth co-stress",])
wa$Basin <- as.factor(wa$Basin)

# make color lists
colors <- c('#4D4D4D', "#AEAEAE", "#E6E6E6")
colors_2 <- c('#55C667', "#97DFFC", "#4A8FE7") # for colorful figures, if you want

# generate the plot with code below (used base R here)
par(mfrow=c(1,2))

par(mar = c(7, 5, 5, 5) + 0.3)   # P stress            
boxplot(wa$value ~ wa$Basin, main="Putative Fe-P co-stress markers", xlab = "", ylab = "Putative Fe-P co-stress marker protein signal", cex.lab=2, cex.axis=1.5, cex.main=2, cex.sub=1.6, col=colors)
abline(h=1, lty=2)
```

# figure s3
```{r}
# define our marker proteins
tricho_markers <- tricho %>%  
  mutate(Category = if_else(Tery_ID == "Tery_1667" | Tery_ID == "Tery_3377" |
                            Tery_ID == "Tery_1666" | Tery_ID == "Tery_1687",
                            "Iron_Stress",
                    if_else(Tery_ID == "Tery_5003" | Tery_ID ==
                            "Tery_4993" | Tery_ID == "Tery_3845" | 
                            Tery_ID == "Tery_3534" |  Tery_ID == "Tery_0398"
                           , "Phosphorus_Stress", 
                    if_else(Tery_ID == "Tery_4136"
                           , "NifH", 
                           NA))))

# iron stress markers and NifH corr df
corr_fe <- tricho_markers %>%
  filter(!is.na(Category)) %>% 
  filter(!Category == "Phosphorus_Stress") %>% # keep only iron marker stress and NifH
  dplyr::select(c(2,4:29)) %>% # keep only ID column and relative abundance data
  remove_rownames() %>% 
  column_to_rownames("Tery_ID") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  dplyr::rename(Fld1 = Tery_1666, IsiA = Tery_1667, FbaA = Tery_1687, IdiA = Tery_3377, NifH = Tery_4136) %>% # rename columns by protein name
  relocate(c(FbaA, IsiA), .before = Fld1) %>% # relocate columns to reflect clustering, but arrange NifH on the right 
  replace(is.na(.), 0) # replace any NA with 0

# do the corr
res2_fe_nifh <- rcorr(as.matrix(corr_fe)) # do pearson corrleation
p.mat_fe_nifh<- cor.mtest(corr_fe) # produce significance for each comparison

# color labels
LabelColFeNifH = c(rep("#CF4D6F",4), "#758BFD")

# phosphorus stress markers and NifH corr df (same thing as above, but for p)
corr_p <- tricho_markers %>%
  filter(!is.na(Category)) %>%
  filter(!Category == "Iron_Stress") %>%
  dplyr::select(c(2,4:29)) %>%
  remove_rownames() %>%
  column_to_rownames("Tery_ID") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  dplyr::rename(SqdB = Tery_0398, SphX = Tery_3534, PhoX = Tery_3845, PhnD = Tery_4993, PhnM = Tery_5003, NifH = Tery_4136) %>%
    replace(is.na(.), 0)

# do the corr
res2_p_nifh <- rcorr(as.matrix(corr_p))
p.mat_p_nifh<- cor.mtest(corr_p)

# color labels
LabelColPNifH = c(rep("#00C49A",5), "#758BFD")

# plot the pearson correlations
par(mfrow=c(2,1))

corrplot(res2_fe_nifh$r, type="upper", order="original", method="color", sig.level = 0.05 , cl.pos="b", vertical=TRUE, tl.cex=2, cl.cex=1.5, tl.col=LabelColFeNifH, p.mat=p.mat_fe_nifh$p, insig="label_sig")

corrplot(res2_p_nifh$r, type="upper", order="hclust", method="color", sig.level = 0.05 , cl.pos="b", vertical=TRUE, tl.cex=2, cl.cex=1.5, tl.col=LabelColPNifH, p.mat=p.mat_p_nifh$p, insig="label_sig")

```

# statistical code examples
```{r}
# KS test for co-stress protein set using df generated above-- this tests the significance of the full co-stress set between regions
ks.test(costress_norm_2$Norm_N_Atl1, costress_norm_2$Norm_S_Pac1) # p = 0.1324
ks.test(costress_norm_2$Norm_S_Pac1, costress_norm_2$Norm_N_Pac1) # p = 0.05602
ks.test(costress_norm_2$Norm_N_Pac1, costress_norm_2$Norm_N_Atl1) # p = 0.1324

# Kruskal-Wallis and Pairwise Wilcoxon code on to test significance of individual putative Fe-P co-stress proteins across regions (using df generated above)
basins <- c(rep("North Atlantic", 14), rep("North Pacific",9), rep("South Pacific", 3)) # make a basins list

test_df <- costress_norm_1 %>%
  mutate(Uniprot = str_extract(Annot, "(?<=_)[^_]+(?=_)")) %>%
  dplyr::select(c(4:29,64)) %>%
  remove_rownames() %>%
  column_to_rownames("Uniprot") %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  mutate(Basins = basins) %>%
  replace(is.na(.), 0)

colnames(test_df) # list column names

# example for putative Fe-P co-stress protein
kruskal.test(Q116X2 ~ Basins, data = test_df)
pairwise.wilcox.test(test_df$Q116X2, test_df$Basins,
                 p.adjust.method = "BH")

# Kruskal wallis and post-hoc pairwise Wilcoxon test: example for testing DIP across regions (from Table S1, read in earlier)
kruskal.test(DIP_nM ~ Ocean_region, data = env)
pairwise.wilcox.test(env$DIP_nM, env$Ocean_region,
                 p.adjust.method = "BH")
```
